The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-11-13 Manuscript_draft/Draft.Rmd
small changes

Here, we hypothesized that:
(1) logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated. This decrease will be compensated by addition of slash which provides labile organic matter, increasing the heterotrophic respiration. 
(2) soil respiration would be more dependent on microclimatic variables (soil temperature and moisture), in the management disturbance treatments where the vegetation was removed and where the litter layer (slash) was removed, by depletion of labile organic matter. Likewise, the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
(4) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.

@@

# Discussion

## Role of temperature and moisture in decomposition
Our model showed that soil temperature and moisture were able to explained more than 50% of variance in soil respiration. Because phenologial patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling. The seasonal scaling although apparently differentiated slightly between sites, it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration. It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%). This is probably related to higher photosynthetic activity variation of Quercus pubescens growing in a Mediterranean climate in France. This would indicate higher changes in root respiration and proliferation that would in turn affect microbial communities. Because in the three study sites there were deciduous Quercus species, another explanation is that soil respiration could have been also affected by the deciduousness of the understorey vegetation (Curiel-Yuste 2004).

## Forest management effect on activation energy and sensitivity to temperature
Our results confirmed the hypothesis 2, since addition of slash decreased the activation energy probably related to a reduction in chemical recalcitrance. This supports the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999). We also observed a positive correlation between Q10 and activation energy agreeing with Davidson & Janssens, 2006 and Sierra, 2012 that the higher the activation energy the higher is the relative temperature sensitivity.

In line with our hypothesis x, we have observed a general negative trend of activation energy with time in all management disturbance treatments and all sites. This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014). Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases of activation energies in the plots where removal of slash had an effect. Additionally, we cannot rule out the fact that other mechanisms could have contributed to low activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022). In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit while in disturbed plots was always descendent. Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022). This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance (Levegre et al 2014).

## Role of edaphic, ecological and microbiological variables in temperature sensitivity
The Q10 values obtained in our study were comparable to the Q10 values measured in the field by other studies (ref). Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0. Fierer et al. (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment.

Hydrologic leaching of dissolved and particulate carbon compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low carbon content in Romania (sandy soil with low retention capacity). Longer time of frozen soils could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), given higher values of Q10 not related to recalcitrance.

>>>>

Here, we hypothesized that:
(1) logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated. This decrease will be compensated by addition of slash which provides labile organic matter, increasing the heterotrophic respiration. 
(2) soil respiration would be more dependent on microclimatic variables (soil temperature and moisture), in the management disturbance treatments where the vegetation was removed and where the litter layer (slash) was removed, by depletion of labile organic matter. Likewise, the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
(3) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.

@@

# Discussion
General paragraph

Hydrologic leaching of dissolved and particulate carbon compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low carbon content in Romania (sandy soil with low retention capacity).

### Role of temperature and moisture in decomposition (model description)
Our model showed that soil temperature and moisture were able to explained more than 50% of variance in soil respiration. Because phenologial patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling. The seasonal scaling although apparently differentiated slightly between sites, it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration. It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%). This is probably related to higher photosynthetic activity variation of Quercus pubescens growing in a Mediterranean climate in France. This would indicate higher changes in root respiration and proliferation that would in turn affect microbial communities. Because in the three study sites there were deciduous Quercus species, another explanation is that soil respiration could have been also affected by the deciduousness of the understorey vegetation (Curiel-Yuste 2004).

## Forest management effect on activation energy and sensitivity to temperature
Generally, our results confirmed the hypothesis 1, since slash addition increased the respiration rates compared to the treatments where the slash was removed. Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (ref). These results indicate that increases in heterotrophic respiration due to due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al. 2015). Moreover, the strength of the response was also dependent on the forest, being more evident in Spain and France where the tree density was much higher than Romania. In Romania such a low density of trees together with a dense understory of grasses in all treatments might explain the lack of response and indicate high degree of resilience to perturbations.

Our results also confirmed the hypothesis 2, since addition of slash decreased the activation energy of SOM decomposition. We argue that this result is probably related to a reduction in chemical recalcitrance, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999). Moreover, the activation energy correlated positively with the Q10, agreeing with Davidson & Janssens, 2006 and Sierra, 2012 that the higher the activation energy the higher is the relative temperature sensitivity. We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites. This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014). Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases in activation energies in the plots where removal of slash had an effect. Additionally, we cannot rule out the fact that other mechanisms could have contributed to low down activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022). In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit, while in disturbed plots it was always descendant. Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022). This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance (Levegre et al 2014).

## Role of edaphic, ecological and microbiological variables in temperature sensitivity
The Q10 values obtained in our study were comparable to the Q10 values measured in the field by other studies (ref). Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0. Fierer et al. (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment. Still, we observed a large variation of Q10 depending on the forest, being the highest in Romania. This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017). The reason could be related to longer time of frozen soils that could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.
